Finiteness, Cardinality, and Combinatorics in Homotopy Type Theory

### 2020-01-05 paper.tex
rm theroems, definitions, lemmas

\subsubsection{Practical Uses of Finiteness}
Proofs of finiteness have well-known practical applications in
constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
In section~\ref{practical}, we build a library which exploits these uses in
Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
complex proofs over finite types.
We frame this in terms of the principle of omniscience for finite types.
Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
go further than the usual examples of this kind of proof automation: as well as
proving properties about functions, we can synthesise functions whole-cloth from
their desired properties.
Through the unified interface for finite and countable types, we can
reuse the automation machinery for \emph{partial} proof search over infinite
search spaces.
Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
prove isomorphisms between the curried and uncurried forms of \(n\)-ary
dependent functions.
\section{Split Enumerability} \label{split-enumerable}
We will start with the simplest definition of finiteness: we say a set is
enumerable if there is a list of its elements which contains every element in
the set.
More formally:
\begin{definition}[Split Enumerable Set]
  \begin{equation}
    \mathcal{E}!(A) = \sum_{(\mathit{xs} : \textbf{List}(A))} , \prod_{(x : A)} , x \in xs
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
\end{definition}

@@

\begin{definition}[List Membership] \label{list-membership}
  \begin{equation}
    x \in \mathit{xs} = \sum_{(i : \mathbf{Fin}(\text{length}(\mathit{xs})))} , \mathit{xs} \ind i \equiv x
  \end{equation}
  Here, \(\ind\) is a safe indexing function.
  In other words, to say that some item \(x\) is in some list \(\mathit{xs}\),
  we say that there is an index into \(\mathit{xs}\) which selects an element of
  \(\mathit{xs}\) which is equivalent to \(x\).
\end{definition}
\begin{definition}[\mathbf{Fin}]
  \(\mathbf{Fin}\)(n) is the type of natural numbers smaller than \(n\).
  It is defined inductively, as follows:
  \begin{equation}
    \begin{aligned}
      \mathbf{Fin}(0) &= \bot \\
      \mathbf{Fin}(n + 1) &= 1 + \mathbf{Fin}(n)
    \end{aligned}
  \end{equation}
\end{definition}

Clearly, this definition of list membership tells us not just that an item is in
a list, but also its \emph{position} in that list.
That piece of information gives use decidable equality, restricting the
predicate to discrete sets.
In section~\ref{manifest-enumerability}, we will show how to hide the position,
giving us a variant of the predicate which allows for non-set finite types.
\subsection{Containers}
One last part is missing from the finiteness definition above: the definition of
\(\mathbf{List}\).
The standard definition is as follows:
\begin{definition}[Lists as an Inductive Type]
  \begin{equation}
    \begin{aligned}
      \mathbf{List}(A) &= \\
      & \left| \; \text{nil} : \mathbf{List}(A)   \right; \\
      & \left| \; \text{cons} : A \rightarrow \mathbf{List}(A) \rightarrow \mathbf{List}(A) \right;
    \end{aligned}
  \end{equation}
\end{definition}

However, as we will see, it will be occasionally more convenient to work with an
equivalent type defined as a \emph{container}.
\begin{definition}[Container] \label{container-def}
  A container \cite{abbottContainersConstructingStrictly2005} is a dependent
  pair of shapes, and positions indexed by those shapes. 
  \begin{equation}
    \mathbf{Container} = \sum_{(\text{shape} : \text{Type})} , \text{shape} \rightarrow \text{Type}
  \end{equation}
  We ``interpret'' a container into a pair of a specific shape, and a function
  which goes from positions to contents.
  \begin{equation}
    \llbracket S , P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
  \end{equation}
  Membership of a container can be defined like so:
  \begin{equation}
    x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
  \end{equation}
\end{definition}

@@

\begin{lemma}
  The two definitions of lists---the traditional inductive type, and the
  container---are equivalent.
  \begin{equation}
    \mathbf{List} \simeq \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
  \end{equation}
\end{lemma}
\begin{proof}
\todo[inline]{Agda code for proof?}
\end{proof}

@@

\begin{theorem} \label{split-enum-surj}
  Split enumerability is equivalent to a split surjection from a finite prefix
  of the natural numbers.
  \begin{equation}
    \mathcal{E}!(A) \simeq \sum_{(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{theorem}
\begin{proof}
  This proof is the composition of two intermediate equivalences: lemmas
  \ref{container-split-enum-equiv} and \ref{container-split-enum-surj}.
  This first of these proves that our definition of finiteness is equivalent to
  another predicate which uses lists defined as a \emph{container}
  (definition~\ref{container-def}).
  The second of these shows that the container-based finiteness predicate is
  equivalent to a split surjection from some \(\mathbf{Fin}\).
\end{proof}

Our interest in the container-based finiteness predicate comes from the fact
that it is very straightforward to prove it equivalent to a split surjection
from some \(\mathbf{Fin}\).
\begin{lemma} \label{container-split-enum-surj}
  Container-defined split enumerability is equivalent to a split surjection from
  some \(\mathbf{Fin}\).
  \begin{equation}
    \mathcal{E}!(A) \simeq \sum_{(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{lemma}
\begin{proof}
  The proof is surprisingly short: after sufficient inlining, it emerges that
  our goal is simply a reassociation.
  \ExecuteMetaData[agda/Cardinality/Container/Finite.tex]{split-surj}
\end{proof}
\subsection{Decidable Equality}
\begin{lemma} \label{split-enum-discrete}
  Any split enumerable type has decidable equality (is discrete).
\end{lemma}
\begin{proof}
  We use a corollary that if there is a split-surjection from \(A\) to \(B\),
  and \(A\) is discrete, then \(B\) is also discrete.
\end{proof}
\begin{lemma}
  Any split enumerable type is a set.
\end{lemma}
\begin{proof}
  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, since split
  enumerable types have decidable equality
  (proposition~\ref{split-enum-discrete}), they are sets.
\end{proof}
\subsection{Closure}
In this section we will prove closure under various operations for split
enumerable sets.
We are working towards a topos proof, which requires us to prove closure under
a variety of operations: for now, we only have enough machinery to demonstrate
the semiring operations, and dependent sums.
in order to show closure under exponentials (function arrows), we will need an
equivalence with \(\mathbf{Fin}\), which will be provided in
section~\ref{manifest-bishop}.
\begin{lemma}
  \(\bot\) is split enumerable.
\end{lemma}
\begin{proof}
  The support list for \(\bot\) is the empty list.
  The cover proof is present via the principle of explosion.
\end{proof}
\begin{lemma}
  \(\top\) is split enumerable.
\end{lemma}
\begin{proof}
  The support list for \(\top\) is the singleton list containing \(\text{tt}\).
  The cover proof is a constant function which returns an index into the first
  item in the list.
\end{proof}
Both of those cases were relatively simple.
Next, we will look into how to combine proofs of split enumerability.
\begin{theorem}
  Split-enumerability is closed under \(\sum\).
\end{theorem}
\begin{proof}
  Let \(E_A\) be a proof of split enumerability for some type \(A\), and \(E_U\)
  be a function of the type:
  \begin{equation}
    E_U : \prod_{(x : A)} , \mathcal{E}!(U(x))
  \end{equation}
  In other words, a function which returns a proof of split enumerability for
  each member of the family \(U\).

@@

  The cover proof follows in a simple, if tedious, way.
\end{proof}
\begin{lemma}
  Split-enumerability is closed under Cartesian product (non-dependent product).
\end{lemma} 
\begin{proof}
  Since non-dependent products are simply a special case of the \(\sum\) type,
  we can reuse the proof of closure over \(\sum\) here.
\end{proof}
\begin{lemma}
  Split-enumerability is closed under disjoint union (non-dependent sum).
\end{lemma}
\begin{proof}
  Again, this can be derived from closure of the \(\sum\) type.
  Disjoint union between two types \(A\) and \(B\) can be represented by the
  following type:
  \begin{equation}
    A + B = \sum_{(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
  \end{equation}
  Then, since all of \(\mathbf{Bool}\), \(A\), and \(B\) are split enumerable,
  the type \(A + B\) is split enumerable.
\end{proof}
\section{Manifest Bishop Finiteness} \label{manifest-bishop}
\begin{definition}[Manifest Bishop Finiteness]
  \begin{equation}
    \mathcal{B}(A) = \sum_{(\mathit{xs} : \textbf{List}(A))} , \prod_{(x : A)} , x \inunique xs
  \end{equation}
\end{definition}
The only difference between this predicate and split enumerability is the list
membership term: we use \(\inunique\) here, where \(x \inunique \mathit{xs}\) is
to be read as ``\(x\) occurs exactly once in \(\mathit{xs}\)''.
It is defined in terms of our previous list membership term
(definition~\ref{list-membership}).
\begin{definition}[Unique List Membership]
  We say an item \(x\) is ``uniquely in'' some list \(\mathit{xs}\) if its
  membership in that list is a \emph{contraction}; i.e. its membership proof
  exists, and all such proofs are equal.
  \begin{equation}
    x \inunique \mathit{xs} = \text{isContr}(x \in \mathit{xs})
  \end{equation}
\end{definition}

A nice consequence of prohibiting duplicates is that now the length of the
support list is the same as the cardinality of the set.
\subsection{Equivalence}
While manifest Bishop finiteness may seem stricter than split enumerability, we
will see soon that it is actually equivalent in strength.
Why study it separately to split enumerability at all, then?
What purpose does it serve on our road to Kuratowski finiteness?
Those questions are answered by the following theorem:
\begin{lemma} \label{bishop-equiv}
  A proof of manifest Bishop finiteness is equivalent to an equivalence with a
  finite prefix of the natural numbers.
  \begin{equation}
    \mathcal{B}(A) \simeq \sum_{(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{lemma}
\begin{proof}
  There are many equivalent definitions of equivalence in HoTT.
  Here we take the version preferred in the Cubical Agda library: contractible
  maps \cite{hottbook}.
  Because of the parallels between contractible maps and split surjections,
  the proof proceeds much the same as \ref{split-enum-surj}.
  In other words, we can define a container-based version of bishop finiteness
  which is a reassociation of a contractible map, and then show that the
  container-based predicate is equivalent to the list-based.
\end{proof}
\subsection{Relationship to Split Enumerability}
We now show that manifest Bishop finiteness has equal strength to split
enumerability.
\begin{lemma}
  Any manifest Bishop finite set is split enumerable.
\end{lemma}
\begin{proof}
  The support set carries over simply, and the cover proof can be taken from the
  first component of the cover proof from the proof of manifest Bishop
  finiteness.
  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{bishop-to-split-enum}
\end{proof}
\begin{theorem} \label{split-enum-to-manifest-bishop}
  Any split enumerable set is manifest Bishop finite.
\end{theorem}
\begin{proof}
  Let \(E\) be a proof of split enumerability for some set \(A\).
  From proposition~\ref{split-enum-discrete} we can derive decidable equality on
  \(A\), and using this we can define a function \(\text{uniques}\) which filters
  out duplicates from lists of \(A\)s.
  \begin{equation}
    \text{uniques} : \textbf{List}(A) \rightarrow \textbf{List}(A)
  \end{equation}
  This gives us our support list.

  It suffices now to prove the following:
  \begin{equation}
    \prod_{(x : A)} , \prod_{(\mathit{xs} : \textbf{List}(A))} , x \in \mathit{xs} \rightarrow x \inunique \text{uniques}(\mathit{xs})
  \end{equation}
  And from that we can generate our cover proof.
\end{proof}
Note that while manifest Bishop finiteness as split enumerability are equivalent
in ``strength'', the two types are not equivalent.
In particular, there are infinitely many inhabitants of \(\mathcal{E}!\), while
for a type \(A\) with \(n\) inhabitants, there are only \(n!\) inhabitants of
\(\mathcal{B}(A)\).
\subsection{Closure}
Proving equal strength of split enumerability and manifest Bishop finiteness
allows us to carry all of the previous proofs of closure over to manifest Bishop
finite sets (and vice-versa).
Missing from our previous proofs was a proof of closure of functions.
We remedy that here.
\begin{theorem}
  Manifest bishop finiteness is closed over dependent functions (\(\prod\)-types).

  Formally, given a type \(A\) and a type family \(U\) on \(A\), when \(A\) is
  manifestly Bishop finite:
  \begin{equation}
    \mathcal{B}\left( A \right)
  \end{equation}
  And \(U\) is manifestly Bishop finite over all points of \(A\):
  \begin{equation}
    \prod_{(x : A)} , \mathcal{B}\left( U(x) \right)
  \end{equation}
  Then we have the following:
  \begin{equation}
    \mathcal{B}\left(\prod_{(x : A)} , U(x)\right)
  \end{equation}
\end{theorem}
\begin{proof}
  This proof is essentially the composition of two transport operations, made
  available to us via univalence.

@@

  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{pi-clos}
\end{proof}
\section{Manifest Enumerability} \label{manifest-enumerability}
As we said in section~\ref{split-enumerable}, we will now retrieve a distinction
between enumerability and manifest Bishop finiteness.
We do so by propositionally truncating the membership proof:
\begin{definition}[Manifest Enumerability]
  \begin{equation}
    \mathcal{E}(A) = \sum_{(\mathit{xs} : \mathbf{List}(A))} , \prod_{(x : A)} , \lVert x \in \mathit{xs} \rVert
  \end{equation}
\end{definition}
\subsection{Higher Homotopy Levels}
By hiding the position, we have essentially removed the ``decidable'' component
from split enumerability.
Our predicate now becomes general enough to work with non-sets: we will show
here that the circle is manifestly enumerable.
\begin{theorem}
  The circle \(S^1\) is manifestly enumerable.
\end{theorem}
\begin{proof}
  As the cover proof is a truncated proposition, we need only consider the point
  constructors, making this poof the same as the proof of split enumerability on
  \(\top\).

@@

  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{circ-inst}
\end{proof}
\subsection{Surjections}
This predicates relation to surjectivity is much the same as split
enumerability's relation to \emph{split} surjectivity.
Deriving a surjection from \(\mathbf{Fin}\) is straightforward:
\begin{lemma}
  A proof of manifest enumerability is equivalent to a surjection from a finite
  prefix of the natural numbers.
\end{lemma}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}
\subsection{Relation to Split Enumerability}
As split enumerability is stronger than manifest enumerability, we can convert
from one to the other easily.
\begin{lemma}
  Any split enumerable type is also manifestly enumerable.
\end{lemma}
\begin{proof}
  The proof carries over via truncation of the cover proof.
\end{proof}
The main missing piece from manifest enumerability is decidable equality: by
truncating the membership proof, we have removed the ability to distinguish
(decidably) between members of the type.
Indeed, resupplying decidable equality is precisely sufficient to recover split
enumerability.

\begin{theorem} \label{manifest-enum-to-split}
  A manifestly enumerable type with decidable equality is split enumerable.
\end{theorem}
\begin{proof}
  The support list stays the same between both enumerability proofs.

@@

\begin{lemma} \label{fin-inj}
  \(\textbf{Fin}\) is injective.
\end{lemma}
\begin{proof}
  Let \(n\) and \(m\) be natural numbers.
  We want to derive a proof of \(n \equiv m\) from \(\textbf{Fin}(n) \equiv
  \textbf{Fin}(m)\).

@@

Another property which does not rely on an internal order is decidable equality.
\begin{theorem} \label{cardinal-finite-discrete}
  Any cardinal-finite set has decidable equality.
\end{theorem}
\begin{proof}
  We first show that decidable equality is a proposition itself.
  We know that if a type \(A\) is a proposition, then the decision over that
  type is also a proposition.
  Then, via Hedberg's theorem, we know that any type with decidable equality is
  a set, meaning that paths in that type are themselves propositions.
  Therefore we can derive that a decision of equality on elements with decidable
  equality is a proposition, and by function extensionality we see that
  decidable equality is itself a proposition.

@@

\subsection{Membership}
We can define membership of a Kuratowski set using these eliminators.
First we define the \(\Diamond\) type, described in the appendix.
\begin{equation}
  \Diamond P = \left\{ \begin{aligned}
    \phi &= \sum_{(x : \text{Type})} , \text{isProp}(x) \\
    \text{trunc}^\phi &= \text{isSetHProp} \\
    []^\phi &= \bot , \text{isProp}\bot \\
    x \dblcolon^\phi \mathit{xs} &= \lVert P(x) \uplus \text{fst}(\mathit{xs}) \rVert , \text{squash} \\
    \text{com}^\phi &= \dots \\
    \text{dup}^\phi &= \dots
  \end{aligned} \right.
\end{equation}
We can define membership then in terms of this.
\begin{definition}[Kuratowski Membership]
  \begin{equation}
    x \in \mathit{xs} = \Diamond (\equiv x) \mathit{xs}
  \end{equation}
\end{definition}

>>>>

\subsubsection{Practical Uses of Finiteness}
Proofs of finiteness have well-known practical applications in
constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
In section~\ref{practical}, we build a library which exploits these uses in
Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
complex proofs over finite types.
We frame this in terms of the principle of omniscience for finite types.
Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
go further than the usual examples of this kind of proof automation: as well as
proving properties about functions, we can synthesise functions whole-cloth from
their desired properties.
Through the unified interface for finite and countable types, we can
reuse the automation machinery for \emph{partial} proof search over infinite
search spaces.
Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
prove isomorphisms between the curried and uncurried forms of \(n\)-ary
dependent functions.
\section{Split Enumerability} \label{split-enumerable}
We will start with the simplest definition of finiteness: we say a set is
enumerable if there is a list of its elements which contains every element in
the set.
More formally:
\begin{rm-definition}[Split Enumerable Set]
  \begin{equation}
    \mathcal{E}!(A) = \sum_{(\mathit{xs} : \textbf{List}(A))} , \prod_{(x : A)} , x \in xs
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
\end{rm-definition}

@@

\begin{rm-definition}[List Membership] \label{list-membership}
  \begin{equation}
    x \in \mathit{xs} = \sum_{(i : \mathbf{Fin}(\text{length}(\mathit{xs})))} , \mathit{xs} \ind i \equiv x
  \end{equation}
  Here, \(\ind\) is a safe indexing function.
  In other words, to say that some item \(x\) is in some list \(\mathit{xs}\),
  we say that there is an index into \(\mathit{xs}\) which selects an element of
  \(\mathit{xs}\) which is equivalent to \(x\).
\end{rm-definition}
\begin{rm-definition}[\mathbf{Fin}]
  \(\mathbf{Fin}\)(n) is the type of natural numbers smaller than \(n\).
  It is defined inductively, as follows:
  \begin{equation}
    \begin{aligned}
      \mathbf{Fin}(0) &= \bot \\
      \mathbf{Fin}(n + 1) &= 1 + \mathbf{Fin}(n)
    \end{aligned}
  \end{equation}
\end{rm-definition}

Clearly, this definition of list membership tells us not just that an item is in
a list, but also its \emph{position} in that list.
That piece of information gives use decidable equality, restricting the
predicate to discrete sets.
In section~\ref{manifest-enumerability}, we will show how to hide the position,
giving us a variant of the predicate which allows for non-set finite types.
\subsection{Containers}
One last part is missing from the finiteness definition above: the definition of
\(\mathbf{List}\).
The standard definition is as follows:
\begin{rm-definition}[Lists as an Inductive Type]
  \begin{equation}
    \begin{aligned}
      \mathbf{List}(A) &= \\
      & \left| \; \text{nil} : \mathbf{List}(A)   \right; \\
      & \left| \; \text{cons} : A \rightarrow \mathbf{List}(A) \rightarrow \mathbf{List}(A) \right;
    \end{aligned}
  \end{equation}
\end{rm-definition}

However, as we will see, it will be occasionally more convenient to work with an
equivalent type defined as a \emph{container}.
\begin{rm-definition}[Container] \label{container-def}
  A container \cite{abbottContainersConstructingStrictly2005} is a dependent
  pair of shapes, and positions indexed by those shapes. 
  \begin{equation}
    \mathbf{Container} = \sum_{(\text{shape} : \text{Type})} , \text{shape} \rightarrow \text{Type}
  \end{equation}
  We ``interpret'' a container into a pair of a specific shape, and a function
  which goes from positions to contents.
  \begin{equation}
    \llbracket S , P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
  \end{equation}
  Membership of a container can be defined like so:
  \begin{equation}
    x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
  \end{equation}
\end{rm-definition}

@@

\begin{rm-lemma}
  The two definitions of lists---the traditional inductive type, and the
  container---are equivalent.
  \begin{equation}
    \mathbf{List} \simeq \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
  \end{equation}
\end{rm-lemma}
\begin{proof}
\todo[inline]{Agda code for proof?}
\end{proof}

@@

\begin{rm-theorem} \label{split-enum-surj}
  Split enumerability is equivalent to a split surjection from a finite prefix
  of the natural numbers.
  \begin{equation}
    \mathcal{E}!(A) \simeq \sum_{(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{rm-theorem}
\begin{proof}
  This proof is the composition of two intermediate equivalences: lemmas
  \ref{container-split-enum-equiv} and \ref{container-split-enum-surj}.
  This first of these proves that our definition of finiteness is equivalent to
  another predicate which uses lists defined as a \emph{container}
  (definition~\ref{container-def}).
  The second of these shows that the container-based finiteness predicate is
  equivalent to a split surjection from some \(\mathbf{Fin}\).
\end{proof}

Our interest in the container-based finiteness predicate comes from the fact
that it is very straightforward to prove it equivalent to a split surjection
from some \(\mathbf{Fin}\).
\begin{rm-lemma} \label{container-split-enum-surj}
  Container-defined split enumerability is equivalent to a split surjection from
  some \(\mathbf{Fin}\).
  \begin{equation}
    \mathcal{E}!(A) \simeq \sum_{(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{rm-lemma}
\begin{proof}
  The proof is surprisingly short: after sufficient inlining, it emerges that
  our goal is simply a reassociation.
  \ExecuteMetaData[agda/Cardinality/Container/Finite.tex]{split-surj}
\end{proof}
\subsection{Decidable Equality}
\begin{rm-lemma} \label{split-enum-discrete}
  Any split enumerable type has decidable equality (is discrete).
\end{rm-lemma}
\begin{proof}
  We use a corollary that if there is a split-surjection from \(A\) to \(B\),
  and \(A\) is discrete, then \(B\) is also discrete.
\end{proof}
\begin{rm-lemma}
  Any split enumerable type is a set.
\end{rm-lemma}
\begin{proof}
  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, since split
  enumerable types have decidable equality
  (proposition~\ref{split-enum-discrete}), they are sets.
\end{proof}
\subsection{Closure}
In this section we will prove closure under various operations for split
enumerable sets.
We are working towards a topos proof, which requires us to prove closure under
a variety of operations: for now, we only have enough machinery to demonstrate
the semiring operations, and dependent sums.
in order to show closure under exponentials (function arrows), we will need an
equivalence with \(\mathbf{Fin}\), which will be provided in
section~\ref{manifest-bishop}.
\begin{rm-lemma}
  \(\bot\) is split enumerable.
\end{rm-lemma}
\begin{proof}
  The support list for \(\bot\) is the empty list.
  The cover proof is present via the principle of explosion.
\end{proof}
\begin{rm-lemma}
  \(\top\) is split enumerable.
\end{rm-lemma}
\begin{proof}
  The support list for \(\top\) is the singleton list containing \(\text{tt}\).
  The cover proof is a constant function which returns an index into the first
  item in the list.
\end{proof}
Both of those cases were relatively simple.
Next, we will look into how to combine proofs of split enumerability.
\begin{rm-theorem}
  Split-enumerability is closed under \(\sum\).
\end{rm-theorem}
\begin{proof}
  Let \(E_A\) be a proof of split enumerability for some type \(A\), and \(E_U\)
  be a function of the type:
  \begin{equation}
    E_U : \prod_{(x : A)} , \mathcal{E}!(U(x))
  \end{equation}
  In other words, a function which returns a proof of split enumerability for
  each member of the family \(U\).

@@

  The cover proof follows in a simple, if tedious, way.
\end{proof}
\begin{rm-lemma}
  Split-enumerability is closed under Cartesian product (non-dependent product).
\end{rm-lemma} 
\begin{proof}
  Since non-dependent products are simply a special case of the \(\sum\) type,
  we can reuse the proof of closure over \(\sum\) here.
\end{proof}
\begin{rm-lemma}
  Split-enumerability is closed under disjoint union (non-dependent sum).
\end{rm-lemma}
\begin{proof}
  Again, this can be derived from closure of the \(\sum\) type.
  Disjoint union between two types \(A\) and \(B\) can be represented by the
  following type:
  \begin{equation}
    A + B = \sum_{(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
  \end{equation}
  Then, since all of \(\mathbf{Bool}\), \(A\), and \(B\) are split enumerable,
  the type \(A + B\) is split enumerable.
\end{proof}
\section{Manifest Bishop Finiteness} \label{manifest-bishop}
\begin{rm-definition}[Manifest Bishop Finiteness]
  \begin{equation}
    \mathcal{B}(A) = \sum_{(\mathit{xs} : \textbf{List}(A))} , \prod_{(x : A)} , x \inunique xs
  \end{equation}
\end{rm-definition}
The only difference between this predicate and split enumerability is the list
membership term: we use \(\inunique\) here, where \(x \inunique \mathit{xs}\) is
to be read as ``\(x\) occurs exactly once in \(\mathit{xs}\)''.
It is defined in terms of our previous list membership term
(definition~\ref{list-membership}).
\begin{rm-definition}[Unique List Membership]
  We say an item \(x\) is ``uniquely in'' some list \(\mathit{xs}\) if its
  membership in that list is a \emph{contraction}; i.e. its membership proof
  exists, and all such proofs are equal.
  \begin{equation}
    x \inunique \mathit{xs} = \text{isContr}(x \in \mathit{xs})
  \end{equation}
\end{rm-definition}

A nice consequence of prohibiting duplicates is that now the length of the
support list is the same as the cardinality of the set.
\subsection{Equivalence}
While manifest Bishop finiteness may seem stricter than split enumerability, we
will see soon that it is actually equivalent in strength.
Why study it separately to split enumerability at all, then?
What purpose does it serve on our road to Kuratowski finiteness?
Those questions are answered by the following theorem:
\begin{rm-lemma} \label{bishop-equiv}
  A proof of manifest Bishop finiteness is equivalent to an equivalence with a
  finite prefix of the natural numbers.
  \begin{equation}
    \mathcal{B}(A) \simeq \sum_{(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{rm-lemma}
\begin{proof}
  There are many equivalent definitions of equivalence in HoTT.
  Here we take the version preferred in the Cubical Agda library: contractible
  maps \cite{hottbook}.
  Because of the parallels between contractible maps and split surjections,
  the proof proceeds much the same as \ref{split-enum-surj}.
  In other words, we can define a container-based version of bishop finiteness
  which is a reassociation of a contractible map, and then show that the
  container-based predicate is equivalent to the list-based.
\end{proof}
\subsection{Relationship to Split Enumerability}
We now show that manifest Bishop finiteness has equal strength to split
enumerability.
\begin{rm-lemma}
  Any manifest Bishop finite set is split enumerable.
\end{rm-lemma}
\begin{proof}
  The support set carries over simply, and the cover proof can be taken from the
  first component of the cover proof from the proof of manifest Bishop
  finiteness.
  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{bishop-to-split-enum}
\end{proof}
\begin{rm-theorem} \label{split-enum-to-manifest-bishop}
  Any split enumerable set is manifest Bishop finite.
\end{rm-theorem}
\begin{proof}
  Let \(E\) be a proof of split enumerability for some set \(A\).
  From proposition~\ref{split-enum-discrete} we can derive decidable equality on
  \(A\), and using this we can define a function \(\text{uniques}\) which filters
  out duplicates from lists of \(A\)s.
  \begin{equation}
    \text{uniques} : \textbf{List}(A) \rightarrow \textbf{List}(A)
  \end{equation}
  This gives us our support list.

  It suffices now to prove the following:
  \begin{equation}
    \prod_{(x : A)} , \prod_{(\mathit{xs} : \textbf{List}(A))} , x \in \mathit{xs} \rightarrow x \inunique \text{uniques}(\mathit{xs})
  \end{equation}
  And from that we can generate our cover proof.
\end{proof}
Note that while manifest Bishop finiteness as split enumerability are equivalent
in ``strength'', the two types are not equivalent.
In particular, there are infinitely many inhabitants of \(\mathcal{E}!\), while
for a type \(A\) with \(n\) inhabitants, there are only \(n!\) inhabitants of
\(\mathcal{B}(A)\).
\subsection{Closure}
Proving equal strength of split enumerability and manifest Bishop finiteness
allows us to carry all of the previous proofs of closure over to manifest Bishop
finite sets (and vice-versa).
Missing from our previous proofs was a proof of closure of functions.
We remedy that here.
\begin{rm-theorem}
  Manifest bishop finiteness is closed over dependent functions (\(\prod\)-types).

  Formally, given a type \(A\) and a type family \(U\) on \(A\), when \(A\) is
  manifestly Bishop finite:
  \begin{equation}
    \mathcal{B}\left( A \right)
  \end{equation}
  And \(U\) is manifestly Bishop finite over all points of \(A\):
  \begin{equation}
    \prod_{(x : A)} , \mathcal{B}\left( U(x) \right)
  \end{equation}
  Then we have the following:
  \begin{equation}
    \mathcal{B}\left(\prod_{(x : A)} , U(x)\right)
  \end{equation}
\end{rm-theorem}
\begin{proof}
  This proof is essentially the composition of two transport operations, made
  available to us via univalence.

@@

  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{pi-clos}
\end{proof}
\section{Manifest Enumerability} \label{manifest-enumerability}
As we said in section~\ref{split-enumerable}, we will now retrieve a distinction
between enumerability and manifest Bishop finiteness.
We do so by propositionally truncating the membership proof:
\begin{rm-definition}[Manifest Enumerability]
  \begin{equation}
    \mathcal{E}(A) = \sum_{(\mathit{xs} : \mathbf{List}(A))} , \prod_{(x : A)} , \lVert x \in \mathit{xs} \rVert
  \end{equation}
\end{rm-definition}
\subsection{Higher Homotopy Levels}
By hiding the position, we have essentially removed the ``decidable'' component
from split enumerability.
Our predicate now becomes general enough to work with non-sets: we will show
here that the circle is manifestly enumerable.
\begin{rm-theorem}
  The circle \(S^1\) is manifestly enumerable.
\end{rm-theorem}
\begin{proof}
  As the cover proof is a truncated proposition, we need only consider the point
  constructors, making this poof the same as the proof of split enumerability on
  \(\top\).

@@

  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{circ-inst}
\end{proof}
\subsection{Surjections}
This predicates relation to surjectivity is much the same as split
enumerability's relation to \emph{split} surjectivity.
Deriving a surjection from \(\mathbf{Fin}\) is straightforward:
\begin{rm-lemma}
  A proof of manifest enumerability is equivalent to a surjection from a finite
  prefix of the natural numbers.
\end{rm-lemma}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}
\subsection{Relation to Split Enumerability}
As split enumerability is stronger than manifest enumerability, we can convert
from one to the other easily.
\begin{rm-lemma}
  Any split enumerable type is also manifestly enumerable.
\end{rm-lemma}
\begin{proof}
  The proof carries over via truncation of the cover proof.
\end{proof}
The main missing piece from manifest enumerability is decidable equality: by
truncating the membership proof, we have removed the ability to distinguish
(decidably) between members of the type.
Indeed, resupplying decidable equality is precisely sufficient to recover split
enumerability.

\begin{rm-theorem} \label{manifest-enum-to-split}
  A manifestly enumerable type with decidable equality is split enumerable.
\end{rm-theorem}
\begin{proof}
  The support list stays the same between both enumerability proofs.

@@

\begin{rm-lemma} \label{fin-inj}
  \(\textbf{Fin}\) is injective.
\end{rm-lemma}
\begin{proof}
  Let \(n\) and \(m\) be natural numbers.
  We want to derive a proof of \(n \equiv m\) from \(\textbf{Fin}(n) \equiv
  \textbf{Fin}(m)\).

@@

Another property which does not rely on an internal order is decidable equality.
\begin{rm-theorem} \label{cardinal-finite-discrete}
  Any cardinal-finite set has decidable equality.
\end{rm-theorem}
\begin{proof}
  We first show that decidable equality is a proposition itself.
  We know that if a type \(A\) is a proposition, then the decision over that
  type is also a proposition.
  Then, via Hedberg's theorem, we know that any type with decidable equality is
  a set, meaning that paths in that type are themselves propositions.
  Therefore we can derive that a decision of equality on elements with decidable
  equality is a proposition, and by function extensionality we see that
  decidable equality is itself a proposition.

@@

\subsection{Membership}
We can define membership of a Kuratowski set using these eliminators.
First we define the \(\Diamond\) type, described in the appendix.
\begin{equation}
  \Diamond P = \left\{ \begin{aligned}
    \phi &= \sum_{(x : \text{Type})} , \text{isProp}(x) \\
    \text{trunc}^\phi &= \text{isSetHProp} \\
    []^\phi &= \bot , \text{isProp}\bot \\
    x \dblcolon^\phi \mathit{xs} &= \lVert P(x) \uplus \text{fst}(\mathit{xs}) \rVert , \text{squash} \\
    \text{com}^\phi &= \dots \\
    \text{dup}^\phi &= \dots
  \end{aligned} \right.
\end{equation}
We can define membership then in terms of this.
\begin{rm-definition}[Kuratowski Membership]
  \begin{equation}
    x \in \mathit{xs} = \Diamond (\equiv x) \mathit{xs}
  \end{equation}
\end{rm-definition}
